---
title: "Red Wine Quality"
author: "Maja Bedenikovic"
date: "2024-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


data=read.csv('winequality-red.csv')
attach(data)
library(ggplot2)
```

# 0. Intro
In the README file on Github, you can read intro information about the data set and variables in it. Here you will only find explanations of the code and the findings.

# 1. Verifying the type of data and null values

```{r}
str(data, na.rm=TRUE)
```

It shows us that all input variables are numerical (continuous), and quality is int (non-continuous) variable. In total we have 1.599 observations and 12 variables. Even with na.rm = TRUE, we have the same amount of observations.

# 2. About variables - summaries

```{r}
summary(data)
```

When it comes to skewness, we are always looking at mean and median value. If the mean is higher than median, then the distribution of the variable is skewed to the right (meaning that lower values are higher in number than higher values), and the other way around. In order to see what kind of distribution is it (We are always happy with a normal distribution, without extreme outliers), I'm using summaries to see the numerical part, and histogram and boxplot to see the graphic notations of the outliers and the shape of the distribution.

Looking at the numerical part (summary), I'm comparing median and mean values to identify skewness. The only variable which is skewed to the left (meaning that it has more higher values than lower values) is density. The rest of the variables is very borderline, making me think that we will see quite nice normal distributions. But I'm not only looking at median and mean values, but also what is the difference between minimal and maximum value, in order to see what kind of spread of data I might see in the graphics. For example, variables residual sugar, fixed acidity and free sulfur dioxide have very big difference between their minimal and maximum value.

Let's see on the graphs how these distributions really look like, and decide what we can do in case that it shows drastic outliers, and if the outliers are even a bad thing.

# Graphics of variables - identifying distributions and outliers

```{r}
ggplot(data,aes(x=fixed.acidity)) + geom_histogram(fill="#704264",colour="black") # has outliers on the high side
ggplot(data,aes(x=volatile.acidity)) + geom_histogram(fill="#704264",colour="black") # has outliers on the high side
ggplot(data,aes(x=citric.acid)) + geom_histogram(fill="#704264",colour="black") # has almost bimodal distribution, and outliers on the high side
ggplot(data,aes(x=residual.sugar)) + geom_histogram(fill="#704264",colour="black") # has outliers on the high side
ggplot(data,aes(x=chlorides)) + geom_histogram(fill="#704264",colour="black") # has outliers on the high side
ggplot(data,aes(x=free.sulfur.dioxide)) + geom_histogram(fill="#704264",colour="black") #has outliers on the high side
ggplot(data,aes(x=total.sulfur.dioxide)) + geom_histogram(fill="#704264",colour="black") #has outliers on the high side
ggplot(data,aes(x=density)) + geom_histogram(fill="#704264",colour="black") #has almost perfect distribution, nothing needs to be changed
ggplot(data,aes(x=pH)) + geom_histogram(fill="#704264",colour="black") #has almost perfect distribution, but still somewhat of visible outliers on high side
ggplot(data,aes(x=sulphates)) + geom_histogram(fill="#704264",colour="black") #outliers on high side
ggplot(data,aes(x=alcohol)) + geom_histogram(fill="#704264",colour="black") # not definite distribution, with some outliers on the high side
ggplot(data,aes(x=quality)) + geom_histogram(fill="#704264",colour="black") # this is output variable, we need to leave this one as it is, as it will be a Y variable in our regression model later.
```

Above this text, you can see the code which I used to output graphics (histograms), and the comments next to it. 

```{r}
boxplot(fixed.acidity)
```











