---
title: "Red Wine Quality"
author: "Maja Bedenikovic"
date: "2024-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


data=read.csv('winequality-red.csv')
attach(data)
library(ggplot2)
```

# 0. Intro
In the README file on Github, you can read intro information about the dataset and variables in it. Here you will only find explanations of the code and the findings.

# 1. Verifying the type of data and null values

```{r}
str(data, na.rm=TRUE)
```

It shows us that all input variables are numerical (continuous), and quality is int (non-continuous) variable. In total we have 1.599 observations and 12 variables. Even with na.rm = TRUE, we have the same amount of observations.

# 2. About variables - summaries

```{r}
summary(data)
```


```{r}
summary(fixed.acidity)
qplot(fixed.acidity)
```

```{r}
summary(volatile.acidity)
qplot(volatile.acidity)
```

```{r}
summary(citric.acid)
qplot(citric.acid)
```

```{r}
summary(residual.sugar)
qplot(residual.sugar)
```



















